---
title: 'STAT 443: Lab 8'
author: 'Wenxuan Zan (61336194)'
date: "12 March, 2023"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(tseries)
library(zoo)
```

### Question 1
```{r}
data <- read.csv("souvenir.txt", header = FALSE)
colnames(data) <- c("Monthly Sales")
sales_ts <- ts(data, start = c(1987,1), frequency = 12)
training_data <- window(sales_ts,start = c(1987,1), end = c(1993,1))
testing_data <- window(sales_ts,start = c(1993,2), end = c(1993,12))
plot(sales_ts,
     ylab = "Monthly Sales",
     main = "Time Series Plot for Monthly Sales Data")
acf(sales_ts,
    main = "Sample acf Values for Monthly Sales Data")
```

Looking at the time series plot, there is clear seasonal variation and an upward trend , and the variation seems to be increasing with time. So I would suggest a multiplicative model. 

### Question 2
```{r}
HWmodel <- HoltWinters(training_data,seasonal = "multiplicative")
plot(HWmodel)
HWmodel
```

The estimated parameter values from the model are: $$\alpha = 0.375,~\beta = 0.046,~\gamma = 0.452$$

### Question 3
```{r}
plot(fitted(HWmodel),
     main = "Holt-Winters Model Decomposition")

```

### Question 4
```{r}
predicted_sales <- predict(HWmodel,
                           n.ahead = 11, 
                           prediction.interval = TRUE,
                           level = 0.95)
plot(testing_data, type = "b", pch = 19)
lines(predicted_sales[,"fit"], col = "red", type = "b", pch = 19)
lines(predicted_sales[,"upr"], col = "firebrick4", lty = "dashed")
lines(predicted_sales[,"lwr"], col = "firebrick4", lty = "dashed")
legend(1993.2,
       100000,
       legend = c("Observed Value", 
                  "Predicted Value", 
                  "95% Prediction Interval Bound"),
       lty = c("solid","solid","dashed"),
       col = c("black", "red","firebrick4"))
```

All predictions are fairly close to the observed values, and all predicted values fall within the 95% prediction interval. In this regard, the forecast is reasonably accurate. 

### Question 5
```{r}
predicted_sales[1:3,"fit"]
```

The forecast for February, March and April of 1993 are 16196.15, 26440.10, and 19170.49 

### Question 6
```{r}
plot(log(sales_ts),
     ylab = "Monthly Sales",
     main = "Log Transformed Time Series")
```

It appears that a logarithm transformation could make the transformed time series follows an additive model where the seasonal component does not vary with trend.  


