---
title: 'STAT 443: Lab 7'
author: 'Wenxuan Zan (61336194)'
date: "05 March, 2023"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(tseries)
library(zoo)
```

### Question 1
```{r}
data <- read.csv("lab7data.csv",header = TRUE)
annual_ts <- ts(data$Annual, start = c(1919), end = c(2008))
ar1model <- arima(annual_ts, order = c(1,0,0),include.mean = TRUE)
ar1model
```

The fitted model is 
$$X_t - \hat\mu = 0.5843(X_{t-1}-\hat\mu)+Z_t$$
where $\hat\mu = -1.9591$ and $Z_t \sim WN(0,1.265)$

### Question 2
```{r}
acf(annual_ts,
    main = "Correlogram for Annual Minimum Temperature Time Series")
```

We would expect acf values for an AR(1) process to decay exponentially, but the current acf values for the temperature data first show a exponential decay until lag 5 and has acf values exceeds the significant cut-off even at large lags. 

### Question 3
```{r}
diffed_data <- diff(annual_ts, lag = 1, difference = 1)
plot(diffed_data,
     ylab = "Annual Minimum Temperatures",
     main = "Differenced Annual Minimum Temperature")
acf(diffed_data,
    main = "Correlogram for Differenced Annual Minimum Temperature")
pacf(diffed_data,
     main = "Partial-autocorrelation Plot")
```

The series of lag 1 difference now appears to be a MA(1) process as the acf plot shows a clear cut-off at lag 1. 

### Question 4
```{r}
arimamodel <- arima(annual_ts, order = c(0,1,1))
arimamodel
```

The fitted model is $$W_t = -0.7504Z_{t-1} + Z_t$$
where $W_t = X_t - X_{t-1}$ and $Z_t \sim WN(0,1.143)$

### Question 5
```{r}
tsdiag(arimamodel)
```

Most of the standardized residuals fall between the $\pm 2$ range with a few exceeding this range. All acf values fall between the $\pm 2/\sqrt{n}$ range except for lag 0 which is expected. The Ljung-Box statistics are not significant for lag smaller than 9, and are significant for lag 9 and 10. Overall, the model fit is reasonable. 

### Question 6
```{r}
q6_result <- tibble(Model = c("AR(1)", "ARIMA(0,1,1) "),
                    AIC = c(round(ar1model$aic,2),round(arimamodel$aic,2)))
kable(q6_result, caption = "AIC for Competing Models")
```

Looking at the above result, the ARIMA(0,1,1) has a smaller AIC compared to AR(1) model, therefore the ARIMA(0,1,1) model is prefered. 















