---
title: 'STAT 443: Lab 9'
author: 'Wenxuan Zan (61336194)'
date: "20 March, 2023"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(tseries)
library(zoo)
```

```{r}
data("LakeHuron")
training <- window(LakeHuron, start = c(1875), end = c(1969))
testing <- window(LakeHuron, start = c(1970), end = c(1972))
```

### Question 1
```{r}
plot(training,
     ylab = "Lake Level",
     main = "Annual Level of Lake Huron From 1875 to 1969")
acf(training)
pacf(training)
```

Looking at the ACF plot, the values of auto-correlation decrease exponentially and the partial ACF plot shows a cut-off at lag 2. I would suggest a ARMA(2,0) model given the above observations. 

### Question 2
```{r}
model <- arima(training, order = c(2,0,0), include.mean = TRUE)
model
```

Fitted model: 
$$X_t - \hat\mu = 1.0617(X_{t-1} - \hat\mu) - 0.2707(X_{t-2}-\hat\mu)+Z_t;$$
where $\hat\mu = 579.0319$ and $~~Z_t\sim WN(0,0.484)$

### Question 3
```{r}
acf(model$residuals,main = "ACF for Residuals")
tsdiag(model)
```

i) The ACF plot for the residuals indicates no significant autocorrelations after lag 0. 

ii) Almost all standardized residuals fall between the $\pm 2$ range, and the p-values for Ljung-Box statistic are not significant even at lag = 10. 

iii) Given the above two observations, the model fits is good 

### Question 4
```{r}
predicted_levels <- predict(model,n.ahead = 3,prediction.interval = TRUE,level = 0.95)
q4result <- tibble(Time = c("1970", "1971", "1972"),prediction = predicted_levels$pred,
                   lower_bound = prediction - 1.96*predicted_levels$se,
                   upper_bound = prediction + 1.96*predicted_levels$se)
kable(q4result)
```

### Question 5
```{r}
q4result[,"Observed"] <- testing
plot(testing, 
     xlim = c(1970.0, 1972.0),
     ylim = c(576,582),
     type = "b", 
     pch = 19, 
     ylab = "Lake Level")
lines(q4result$prediction, col = "red", type = "b", pch = 19)
lines(q4result$lower_bound, col = "firebrick4", lty = "dashed")
lines(q4result$upper_bound, col = "firebrick4", lty = "dashed")
legend(1970.0,
       578.1,
       legend = c("Observed Value",
                  "Predicted Value",
                   "Prediction Interval"),
       lty = c("solid","solid","dashed"),
       col = c("black", "red","firebrick4"))
```

The prediction interval is wider for predictions that are further into the future. Although all observed values are within the 95% prediction interval, we should notice that the predicted values show a downward trend whereas the observed values showed a upward trend. 




















