---
title: 'STAT 443: Lab 6'
author: 'Wenxuan Zan (61336194)'
date: "20 Feb, 2023"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(tseries)
library(zoo)
```

### Question 1
```{r}
data <- read.csv("TempPG.csv",header = TRUE)
annual_ts <- ts(data$Annual, start = c(1919), end = c(2008))
plot(annual_ts,
     ylab = "Annual Minimum Temperatures",
     main = "Annual Minimum Temperature at Prince George 1919-2008")
acf(annual_ts,
    main = "Correlogram for Annual Minimum Temperature Time Series")
pacf(annual_ts,
     main = "Partial-autocorrelation Plot")
```

i) Looking at the plot titled "Annual Minimum Temperature at Prince George 1919 to 2008", we can see a clear upward trend such that the annual minimum temperature at Prince George from 1919 to 2008 is increasing. Therefore this time series is likely to be non-stationary. 

ii) Looking at the acf plot we observed a slow exponential decay of acf values. This pattern indicates the existence of positive temporal dependence within the data.This pattern is characteristic of a AR model.

iii) Looking at the pacf plot, a sensible cut-off is at lag 1 which suggests a AR(1) model, but the pacf value also spikes at lag 5, and 13, so perhaps a ARMA model is more suitable for the data.

iv) I would suggest a ARMA(1,0) model for the data.

### Question 2
```{r}
model <- arima(annual_ts, order = c(1,0,0), include.mean = TRUE)
model
```

The fitted model is: 

$$X_t + 1.9591 = 0.5843(X_{t-1} + 1.9591) + Z_t;~~~~Z_t \sim WN(0,1.265)$$

### Question 3
```{r}
confint(model)
```

The 95% CI for alpha is $[0.415, 0.754]$ and the 95% CI for $\mu$ is $[-2.510,-1.408]$






